---
title: "Assignment 5:Data Visualization (Incarceration)"
author: "Aisha Abdulkarim"
date: "2024-06-26"
output: html_document
---

### Introduction

As long as African Americans have been here, they have seen high and unprecedented conflict within the United States criminal justice system if not 4 times than their white counterparts. Racial disparities are a problem that spans even through the daily lives of African Americans and other people of color but is highly seen within the Black community. For this assignment, we are going to investigate the population of Black people in Jail across three states, Washington, California, and Texas, from 2000 to 2015. To get a better understanding of which year was there a high incarceration rate as well as identify which state had the highest incarceration. Additionally, we are also going to which gender has the highest incarceration rate. 

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

These will likely be calculated using your DPLYR skills, answering questions such as:

- What is the average value of my variable across all the counties (in the current year)?
- Where is my variable the highest / lowest?
- How much has my variable change over the last N years?

Feel free to calculate and report values that you find relevant. Again, remember th  at the purpose is to think about how these measure of incarceration vary by race.

### The Dataset

Who collected the data?

  - Vera obtained their data from federally collected data through BJS and other various sources for example they used: 
    - Annual Survey of Jails ( ASJ)
    - Census of Jails (COJ)
    - Mortality in Correctional Institutions (MCI)
    - National Prisoner Statistics (NPS)
    - National Corrections Reporting Program (NCRP)
    - The U.S Department of Justice Bureau of Justice Statistics 
    - National Center for Health Statistics (NCHS)

How was the data collected or generated?  

  - They were able to collect the data by going through the BJS Annual Survey of Jails(ASJ), Census of Jails (COJ), and Mortality of Correctional Institutions (MCI) data collections. 

Why was the data collected?  

  - They wanted to give the public access to current and historical data about the jail and prison population not just locally but also across states in the United States.

How many observations (rows) are in your data?  

  - There is a total of 153,811 rows in my data.

How many features (columns) are in the data?

  - There is a total of 121 columns in my data. 

What, if any, ethical questions or questions of power do you need to consider when working with this data?  

  - 
  
What are possible limitations or problems with this data?   (at least 200 words)

  - 

### Trends Over Time Chart

I have included this grouped bar chart to represent my trends over time data. In this graph, I am measuring which state had the highest incarceration levels of Black individuals between the years of 2000 to 2015. The bar chart was the best way to understand what is being compared and evaluated, even if I did a lined chart it would show what I am trying to compare but would make it hard to understand the graph. The importance of this graph is that it allows us to look at the state that had the highest incarceration rate each year and allows us to think of what problem is causing a repetition of such high incarceration rates in one state but the other this is helpful in figuring out what the other states can implement to help lower incarceration rates in their state.


```{r, echo = FALSE, code = readLines("My Chart 1.R")}
```


### Variable Comparison Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show how two different (continuous) variables are related to one another. Again, think carefully about what such a comparison means and what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- If you choose to add a color encoding (not required), you need a legend for your different color and a clear legend title

```{r, echo = FALSE, code = readLines("My Chart 2.R")}
```

### Map

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The last chart that you will create and include will show how a variable is distributed geographically. Again, think carefully about what such a comparison means and what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Your map needs a title
- Your color scale needs a legend with a clear label
- Use a map based coordinate system to set the aspect ratio of your map

```{r, echo = FALSE, code = readLines("My Map.R")}
```


Thoughtfully creates a chart comparing two variables, describes *why* it was designed that way, and what patterns emerge. Chart meets requirements outlined above.