---
title: "Data Visualization (Incarceration)"
author: "Aisha Abdulkarim"
date: "2024-06-26"
output: html_document
---

### Introduction

As long as African Americans have been here, they have seen high and unprecedented conflict within the United States criminal justice system if not 4 times than their white counterparts. Racial disparities are a problem that spans even through the daily lives of African Americans and other people of color but is highly seen within the Black community. For this assignment, we are going to investigate the population of Black people in Jail across three states, Washington, California, and Texas, from 2000 to 2015. To get a better understanding of which year was there a high incarceration rate as well as identify which state had the highest incarceration. Additionally, we are also going to which gender has the highest incarceration rate. 

### Summary Information

In this project, I mainly focused on the years  2000 to 2015 but in this section, I have strictly focused on the black jail population and the year 2005 as that is the year of the highest incarnation rates between the years that I have compared in my charts. I have identified that the average value of black_jail_pop across all counties was **94141.58**. The state with the highest black jail population was **New York, with 7597** and the lowest was **0 in  Alaska**. Knowing that Texas has the highest black population I was surprised that more black people were incarcerated in New York compared to Texas in 2005. I also wanted to investigate which county had the maximum value of black_jail_pop across all states and I found that **Los Angeles, California** had the highest incarceration having incarcerated **13143.92** people.

### The Dataset

This dataset makes an observation of incarceration rates between different states and groups of people within the United States. We are using  Vera and the data they have obtained, they have obtained their data from federally collected data through BJS and multiple other resources such as the **Annual Survey of Jails ( ASJ), Census of Jails (COJ), Mortality in Correctional Institutions (MCI), National Prisoner Statistics (NPS), National Corrections Reporting Program (NCRP), The U.S Department of Justice Bureau of Justice Statistics, National Center for Health Statistics (NCHS). They were able to collect the data by going through the BJS Annual Survey of Jails(ASJ), Census of Jails (COJ), and Mortality of Correctional Institutions (MCI) data collections**. They wanted to give the public access to current and historical data about the jail and prison population not just locally but also across states in the United States. In the data that I am observing there is a total of **153,811 rows** in my data. There are a total of **121 columns** in my data. Some ethical questions that I have when working with this data would be if some of the data they have analyzed from different organizations have the data of the number of people arrested was consent to be added as an individual and if they don't consent would that have an effect of their data, by un-stabilizing it. Some possible limitations with this data that might arise are incomplete data, there are some portions of this data where many columns are empty which might end up giving us errors when making our charts or trying to analyze one section to another. A problem that might arise is that the data might be an overall generalization as it doesn't go in-depth into what crime was committed, which might be a problem when someone is trying to compare the level of crimes being committed in those populations of groups. 

### Trends Over Time Chart

I have included this grouped bar chart to represent my trends over time data. In this graph, I am measuring which state had the highest incarceration levels of Black individuals between the years of 2000 to 2015. The bar chart was the best way to understand what is being compared and evaluated, even if I did a lined chart it would show what I am trying to compare but would make it hard to understand the graph. The importance of this graph is that it allows us to look at which state had the highest incarceration rate each year and allows us to think of what problem is causing a repetition of such high incarceration rates in one state but in another state. This is helpful in figuring out what the other states can implement to help lower incarceration rates in their state. The graph shows that in all the states compared, the state of California has a history of higher incarceration levels of black individuals, particularly in 2005 being the highest between the years I observed. This can also be concluded by how California has the highest population of African Americans and the largest land bases compared to Washington and Texas. 

```{r, echo = FALSE, code = readLines("My Chart 1.R")}
```

  
### Variable Comparison Chart

In this line chart, I decided to compare male jail Admission compared to  Female jail admissions between 2000 and 2015. The primary reason was to examine which gender was more likely to be incarcerated during this period of time and explore the potential reasons behind any observed disparities. The graph illustrate that males consistently experienced higher rates of incarceration compared to females, throughout the analyzed time frame. This graphs along with how men are more subjugated with harshest and stricter criminal charges. Gender roles in crime sentences also play are roll, for example is a women and man commited a harsh crime alot of people would not belife that a women commit said crime. The overrall trend shows that women are in the year 2000, showed a significant incrimination rate compare to all the other years, this makes me question what happened in 2000 that raised the incarceration rates. 

```{r, echo = FALSE, code = readLines("My Chart 2.R")}
```

### Map

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The last chart that you will create and include will show how a variable is distributed geographically. Again, think carefully about what such a comparison means and what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Your map needs a title
- Your color scale needs a legend with a clear label
- Use a map based coordinate system to set the aspect ratio of your map

```{r, echo = FALSE, code = readLines("My Map.R")}
```


